The document title

### 2021-03-15 manuscript/manuscript.Rmd
Update results

FP and accelerometer data were processed using MATLAB (Version 8.3, Mathworks, Natick, USA), as described elsewhere [@Veras2020]. Briefly, both GRF and ACC signals were filtered using a Butterworth fourth-order low-pass filter, with 20 Hz cut-off frequency, and their resultant vector calculated ($r_i = \sqrt{x_i^2 + y_i^2 + z_i^2}$). After that, peak ACC (pACC) values were defined as minimum height of twice the positive ACC average, during any given speed, and separated by 0.4 seconds minimum. GRF and ACC signals were adjusted through the time set by the systems clock, with manual correction based on visual inspection for possible errors, and then synchronized by the maximum cross-correlation coefficient. Peak GRF (pGRF) was determined as the highest value between 0.4 seconds before and after each pACC. The rates of change, LR for the GRF and ACC rate (AR) for the ACC, were computed through a second-order centered derivative from the beginning of the foot contact to the pGRF. The pLR and peak AR (pAR) were defined as the maximum value in the array.

@@

Regression equations were developed to predict the mechanical loading variables based on accelerometry data. In the pGRF prediction equations, body mass and pACC were significant predictors and included in the models. Table 1 shows the regression coefficients, $\mathrm{R^2}$ and accuracy indices from models derived from accelerometers placed at ankle, lower back and hip, and from the resultant vector and its vertical component. In all pGRF prediction models, $\mathrm{R^2}$ ranged from `r min(GRF_accuracy$R2_cond)` to `r max(GRF_accuracy$R2_cond)`, which shows that the predictor variables are able to explain the variability in the outcome varibale in a great extent. Also, MAE ranged between `r myround(min(GRF_accuracy$MAE), 1)` N and `r myround(max(GRF_accuracy$MAE), 1)` N, and the MAPE from `r myround(min(GRF_accuracy$MAPE) * 100, 1)`\% to `r myround(max(GRF_accuracy$MAPE) * 100, 1)`\%, which is a smaller error than previously published equations [@Neugebauer2014; @Fortune2014], and similar to the error from equations previously published by our group, but which included walking only [@Veras2020]. RMSE were also computed and shown to range between `r myround(min(GRF_accuracy$RMSE), 1)` N and `r myround(max(GRF_accuracy$RMSE), 1)` N. Of all the pGRF predictions equations developed, the one based on hip-worn accelerometers and the resultant vector has shown the best performance, with an $\mathrm{R^2}$ of `r hip_GRF_accuracy$R2_cond`, and the lowest MAE (`r myround(hip_GRF_accuracy$MAE, 1)` N), MAPE (`r myround(hip_GRF_accuracy$MAPE * 100, 1)`\%) and RMSE (`r myround(hip_GRF_accuracy$RMSE, 1)` N).

@@

```{r LR_accuracy}
load(here("output", "prediction_models.rda"))
res_LR_accuracy <- purrr::map(
  accuracy_res_LR_models,
  ~ dplyr::mutate(.x, MAPE = as.numeric(MAPE))
) %>%
  purrr::map_dfr(rbind) %>%
  dplyr::mutate(
    placement = c("lower back", "hip"),
    vector = "resultant"
  )
ver_LR_accuracy <- purrr::map(
  accuracy_ver_LR_models,
  ~ dplyr::mutate(.x, MAPE = as.numeric(MAPE))
) %>%
  purrr::map_dfr(rbind) %>%
  dplyr::mutate(
    placement = c("lower back", "hip"),
    vector = "vertical"
  )
LR_accuracy <- rbind(res_LR_accuracy, ver_LR_accuracy)
back_LR_accuracy <- LR_accuracy %>%
  filter(placement == "lower back" & vector == "resultant")
```

Moreover, pLR prediction equations were developed, and these included as predictors the pAR and again body mass. Regression coefficients, $\mathrm{R^2}$ and accuracy indices for these equations are also shown in Table 1. For these equations, the lowest $\mathrm{R^2}$ found was `r min(LR_accuracy$R2_cond)` and the highest was `r max(LR_accuracy$R2_cond)`, reflecting a moderate prediction ability. MAE, MAPE and RMSE values ranged from `r myround(min(LR_accuracy$MAE), 1)` N to `r myround(max(LR_accuracy$MAE), 1)` N, from `r myround(min(LR_accuracy$MAPE) * 100, 1)`\% to `r myround(max(LR_accuracy$MAPE) * 100, 1)`\%, and from `r myround(min(LR_accuracy$RMSE), 1)` N to `r myround(max(LR_accuracy$RMSE), 1)` N, respectively. The results found for the pLR prediction equations were all considerably poorer than those for the pGRF equations. The accuracy indices values were also somewhat worse than previously published pLR prediction equations [@Veras2020; @Fortune2014], although equations from both studies were only validate for walking. The best-performing pLR prediction equation was the one for accelerometers worn at lower back and the resultant vector, with an $\mathrm{R^2}$ of `r back_LR_accuracy$R2_cond` and a MAE, MAPE and RMSE of `r myround(back_LR_accuracy$MAE, 1)` N, `r myround(back_LR_accuracy$MAPE * 100, 1)`\% and `r myround(back_LR_accuracy$RMSE, 1)` N, respectively.

@@

```{r ba_regression}
source(here("code", "04_sub_analyses.R"))
```

Afterwards, Bland-Altman plots were built to visualize the agreement between the actual and predicted pGRF and pLR, for both the resultant vector and its vertical component and all accelerometer placements. A good agreement in a Bland-Altman plot is observed when values tend to cluster around zero, showing no consistent over or underestimation of the actual values. Also, the majority of data points should lie within the limits of agreement, which are determined by \pm 1.96 standard deviation of the difference between actual and predicted values. Bland-Altman plots for the pGRF and pLR resultant vector from hip-worn accelerometers can be seen in Figure 3, and for all accelerometers placements and both resultant and vertical vectors in the Supplemental Figures S3 and S4. Values for all pGRF and pLR prediction equations developed tended to assemble around zero and between the limits of agreement. For the pGRF equations, a tendency to higher dispersion can be observed in the higher magnitudes, which correspond to the running trials, and especially in values derived from ankle-worn accelerometers (Supplemental Figure S3, Panels A and D). Prediction equations for pLR showed a more accentuated dispersion, and again in the higher magnitudes. None of the regression models presented a bias different than zero ($p > 0.05$), demonstrating that these models do not consistently over or underestimate the pGRF and pLR values. Furthermore, linear regressions were applied to identify proportional bias, that is, to test whether pGRF and pLR prediction errors were systematically influenced by their magnitude. The pGRF prediction equation based on ankle-worn accelerometers data for the resultant vector showed a significant proportional bias ($p = `r myround(coefficients(ba_regression_res_GRF$ankle)[2, 4], 2)`$), although with a low magnitude ($\mathrm{R^2} = `r myround(ba_regression_res_GRF$ankle["adj.r.squared"], 2)`$).

>>>>

FP and accelerometer data were processed using MATLAB (Version 2019a, Mathworks, Natick, USA), as described elsewhere [@Veras2020]. Briefly, both GRF and ACC signals were filtered using a Butterworth fourth-order low-pass filter, with 20 Hz cut-off frequency, and their resultant vector calculated ($r_i = \sqrt{x_i^2 + y_i^2 + z_i^2}$). After that, peak ACC (pACC) values were defined as minimum height of twice the positive ACC average, during any given speed, and separated by 0.4 seconds minimum. GRF and ACC signals were adjusted through the time set by the systems clock, with manual correction based on visual inspection for possible errors, and then synchronized by the maximum cross-correlation coefficient. Peak GRF (pGRF) was determined as the highest value between 0.4 seconds before and after each pACC. The rates of change, LR for the GRF and ACC rate (AR) for the ACC, were computed through a second-order centered derivative from the beginning of the foot contact to the pGRF. The pLR and peak AR (pAR) were defined as the maximum value in the array.

@@

Regression equations were developed to predict the mechanical loading variables based on accelerometry data. In the pGRF prediction equations, body mass and pACC were significant predictors and included in the models. Table 1 shows the regression coefficients, $\mathrm{R^2}$ and accuracy indices from models derived from accelerometers placed at ankle, lower back and hip, and from the resultant vector and its vertical component. In all pGRF prediction models, $\mathrm{R^2}$ values were `r max(GRF_accuracy$R2_cond)`, which shows that the predictor variables are able to explain the variability in the outcome varibale to a great extent. Also, MAE ranged between `r myround(min(GRF_accuracy$MAE), 1)` N and `r myround(max(GRF_accuracy$MAE), 1)` N, and the MAPE from `r myround(min(GRF_accuracy$MAPE) * 100, 1)`\% to `r myround(max(GRF_accuracy$MAPE) * 100, 1)`\%, which is a smaller error than previously published equations [@Neugebauer2014; @Fortune2014], and similar to the error from equations previously published by our group, but which included walking only [@Veras2020]. RMSE were also computed and shown to range between `r myround(min(GRF_accuracy$RMSE), 1)` N and `r myround(max(GRF_accuracy$RMSE), 1)` N. Of all the pGRF predictions equations developed, the one based on hip-worn accelerometers and the resultant vector has shown the best performance, with an $\mathrm{R^2}$ of `r hip_GRF_accuracy$R2_cond`, and the lowest MAE (`r myround(hip_GRF_accuracy$MAE, 1)` N), MAPE (`r myround(hip_GRF_accuracy$MAPE * 100, 1)`\%) and RMSE (`r myround(hip_GRF_accuracy$RMSE, 1)` N).

@@

```{r LR_accuracy}
load(here("output", "prediction_models.rda"))
res_LR_accuracy <- purrr::map(
  accuracy_res_LR_models,
  ~ dplyr::mutate(.x, MAPE = as.numeric(MAPE))
) %>%
  purrr::map_dfr(rbind) %>%
  dplyr::mutate(
    placement = c("ankle", "lower back", "hip"),
    vector = "resultant"
  )
ver_LR_accuracy <- purrr::map(
  accuracy_ver_LR_models,
  ~ dplyr::mutate(.x, MAPE = as.numeric(MAPE))
) %>%
  purrr::map_dfr(rbind) %>%
  dplyr::mutate(
    placement = c("ankle", "lower back", "hip"),
    vector = "vertical"
  )
LR_accuracy <- rbind(res_LR_accuracy, ver_LR_accuracy)
back_LR_accuracy <- LR_accuracy %>%
  filter(placement == "lower back" & vector == "resultant")
```

Moreover, pLR prediction equations were developed, and these included as predictors the pAR and again body mass. Regression coefficients, $\mathrm{R^2}$ and accuracy indices for these equations are also shown in Table 1. For these equations, the lowest $\mathrm{R^2}$ found was `r myround(min(LR_accuracy$R2_cond), 2)` and the highest was `r myround(max(LR_accuracy$R2_cond), 2)`, reflecting a moderate to high prediction ability. MAE, MAPE and RMSE values ranged from `r myround(min(LR_accuracy$MAE), 1)` N to `r myround(max(LR_accuracy$MAE), 1)` N, from `r myround(min(LR_accuracy$MAPE) * 100, 1)`\% to `r myround(max(LR_accuracy$MAPE) * 100, 1)`\%, and from `r myround(min(LR_accuracy$RMSE), 1)` N to `r myround(max(LR_accuracy$RMSE), 1)` N, respectively. The results found for the pLR prediction equations were all considerably poorer than those for the pGRF equations. The accuracy indices values were also somewhat worse than previously published pLR prediction equations [@Veras2020; @Fortune2014], although equations from both studies were only validate for walking. The pLR equation with the lower prediction error was the one for accelerometers worn at lower back and the resultant vector, with an $\mathrm{R^2}$ of `r back_LR_accuracy$R2_cond` and a MAE, MAPE and RMSE of `r myround(back_LR_accuracy$MAE, 1)` N, `r myround(back_LR_accuracy$MAPE * 100, 1)`\% and `r myround(back_LR_accuracy$RMSE, 1)` N, respectively.

@@

```{r ba_regression}
source(here("code", "04_sub_analyses.R"))
ba_regression_ankle_R2 <- c(
  ba_regression_res_GRF$ankle["adj.r.squared"],
  ba_regression_ver_GRF$ankle["adj.r.squared"],
  ba_regression_res_LR$ankle["adj.r.squared"],
  ba_regression_ver_LR$ankle["adj.r.squared"]
) %>%
  as_vector()
```

Afterwards, Bland-Altman plots were built to visualize the agreement between the actual and predicted pGRF and pLR, for both the resultant vector and its vertical component and all accelerometer placements. A good agreement in a Bland-Altman plot is observed when values tend to cluster around zero, showing no consistent over or underestimation of the actual values. Also, the majority of data points should lie within the limits of agreement, which are determined by \pm 1.96 standard deviation of the difference between actual and predicted values. Bland-Altman plots for the pGRF and pLR resultant vector from hip-worn accelerometers can be seen in Figure 3, and for all accelerometers placements and both resultant and vertical vectors in the Supplemental Figures S3 and S4. Values for all pGRF and pLR prediction equations developed tended to assemble around zero and between the limits of agreement. For the pGRF equations, a tendency to higher dispersion can be observed in the higher magnitudes, which correspond to the running trials, and especially in values derived from ankle-worn accelerometers (Supplemental Figure S3, Panels A and D). Prediction equations for pLR showed a more accentuated dispersion, and again in the higher magnitudes (Supplemental Figure S4). The pLR prediction model based on data from accelerometers placed at the ankle and the resultant vector displayed a bias different than zero ($p = `r myround(ba_bias_test_res_LR$ankle["p.value"], 2)`$). None of the other models presented a bias different than zero ($p > 0.05$), demonstrating that they do not consistently over or underestimate the pGRF and pLR values. Furthermore, linear regressions were applied to identify proportional bias, that is, to test whether pGRF and pLR prediction errors were systematically influenced by their magnitude. All equations based on ankle-worn accelerometer data, for both outcomes (pGRF and pLR) and vectors (resultant and vertical) showed a significant proportional bias ($p < 0.05$). Despite the statistical significance, these linear regressions presented a low $\mathrm{R^2}$ (highest $\mathrm{R^2} = `r myround(max(ba_regression_ankle_R2), 2)`$).
